---
title: "Assembling PDFs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assembling PDFs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Splitting, merging, reordering and rotating are one operation in zupdf:
copying pages from open files into a new one with `pdf_copy_pages()`. Each
copied page brings the fonts, images and forms it uses, so the result stands
alone.

```{r}
library(zupdf)
src <- pdf_open(system.file("examples", "hello.pdf", package = "zupdf"))
w <- pdf_new()
pdf_copy_pages(w, src, pages = c(2, 1))   # reverse the pages
pdf_copy_pages(w, src, pages = 1, rotate = 90)
out <- pdf_open(pdf_save(w))
pdf_pages(out)[, c("page", "rotate")]
pdf_page_text(out)
pdf_close(out)
```

## Metadata and encryption

`pdf_set_meta()` sets the document information. `pdf_set_encryption()` must
come first, before any page: it fixes the keys.

```{r}
w <- pdf_new()
pdf_set_encryption(w, user = "secret", permissions = c("print", "copy"))
pdf_set_meta(w, title = "Two pages", author = "zupdf")
pdf_copy_pages(w, src)
bytes <- pdf_save(w)
locked <- pdf_open(bytes, password = "secret")
pdf_meta(locked)[c("title", "encryption", "permissions")]
pdf_close(locked)
pdf_close(src)
```

## The qpdf way

The same operations exist under qpdf's names and arguments, working on files:

```{r}
dir <- tempfile()
dir.create(dir)
file.copy(system.file("examples", "hello.pdf", package = "zupdf"), dir)
input <- file.path(dir, "hello.pdf")
basename(pdf_split(input))
pdf_length(pdf_combine(c(input, input)))
```
