---
title: "Switching from pdftools and qpdf"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Switching from pdftools and qpdf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

zupdf provides the functions of pdftools and qpdf that need no renderer,
under their names and with exactly their arguments, so code written for
either runs with `library(zupdf)` in their place:

```{r}
library(zupdf)
path <- system.file("examples", "hello.pdf", package = "zupdf")
pdf_info(path)$pages
pdf_text(path)
pdf_data(path)[[1]]
pdf_length(path)
```

| From | Provided | Not provided |
|---|---|---|
| pdftools | `pdf_info()`, `pdf_text()`, `pdf_data()`, `pdf_fonts()`, `pdf_pagesize()`, `pdf_toc()`, `pdf_attachments()` | `pdf_render_page()`, `pdf_convert()`, `pdf_ocr_text()`, `pdf_ocr_data()`, `poppler_config()` |
| qpdf | `pdf_length()`, `pdf_split()`, `pdf_subset()`, `pdf_combine()`, `pdf_rotate_pages()` | `pdf_compress()`, `pdf_overlay_stamp()` |

What is not provided is not defined at all, so a call fails with "could not
find function" rather than doing something else.

## Where results differ

* `pdf_text()` comes from zupdf's extractor, not poppler's physical layout:
  the same pages and words, single spaces where poppler pads to page
  positions, and no text from annotations (form fields and stamps).
* `pdf_data()`'s word boxes agree with poppler's to within a point or two;
  poppler takes some fonts' heights from elsewhere.
* `pdf_fonts()` reports `file` as `""` for an embedded font and `NA`
  otherwise: zupdf looks up no system fonts.
* `pdf_info()` reports an absent date as `NA`.
* The qpdf functions write through pdfio, so the output's bytes differ from
  qpdf's; its pages, content and order are the same. Outlines and forms are
  not carried over.

## Attaching both

R lets the package attached last win, and says so. Attaching zupdf after
pdftools or qpdf masks their functions of the same name; `pdftools::` and
`qpdf::` still reach theirs.
