---
title: "Inspecting a PDF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inspecting a PDF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

zupdf reads PDF files through the bundled pdfio library. It never renders a
page to pixels; everything else a PDF holds, it can read.

```{r}
library(zupdf)
path <- system.file("examples", "hello.pdf", package = "zupdf")
pdf <- pdf_open(path)
pdf
```

`pdf_open()` reads only the cross-reference table, so opening is cheap; each
function then reads what it needs. A raw vector or a connection works as well
as a path.

## The document

```{r}
str(pdf_meta(pdf))
pdf_pages(pdf)
```

## Text

`pdf_page_text()` follows each page's content stream, maps every character
through its font, and lays the text out in reading order, or in the order the
stream draws it:

```{r}
pdf_page_text(pdf)
pdf_page_text(pdf, pages = 1, layout = "raw")
```

A character no font maps comes back as U+FFFD and is counted in the
`unmapped` attribute.

## Objects and streams

Every object is reachable, as R values: dictionaries are named lists, names
carry the class `pdf_name`, references `pdf_ref`.

```{r}
pdf_objects(pdf)
pdf_object(pdf, 4)
rawToChar(pdf_stream(pdf, 5))
pdf_page_tokens(pdf, 1)
```

## Untrusted files

A PDF is a common carrier of hostile input. `pdf_open()` takes limits on the
input size, the number of objects and pages, the size of any decoded stream,
and the depth of nesting; reaching one is a `zupdf_limit_error`:

```{r}
tryCatch(
  pdf_open(path, max_pages = 1),
  zupdf_limit_error = function(e) conditionMessage(e)
)
pdf_close(pdf)
```
