---
title: "A tour of zupdf"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This article runs through the main things zupdf does, in one session: write a
PDF, read it back, take it apart and put it together again, lock it, and use
the pdftools and qpdf names. Each step has its own guide for the details.

```{r}
library(zupdf)
dir <- tempfile("tour-")
dir.create(dir)
```

## Write a PDF

A writer collects pages. Each page records what is drawn on it, text, paths and
images, and `pdf_page_end()` writes it. `pdf_save()` returns the bytes, or
writes them to a file.

```{r}
w <- pdf_new(media_box = pdf_paper("a4"))
pdf_set_meta(w, title = "Ring of Fire", author = "zupdf")

page <- pdf_page_new(w)
pdf_draw_text(page, 297.6, 780, "Maunga Whau", font = "Helvetica-Bold",
  size = 24, align = "centre")
pdf_draw_text(page, 72, 740, "Heights of a volcano in Auckland, from R's volcano data.",
  font = "Times-Roman", size = 12)

# An image from an R matrix of grey levels between 0 and 1, one row per line
# of pixels: 61 lines of 87 pixels, the shape of the box it is drawn into.
heights <- t(volcano - min(volcano)) / diff(range(volcano))
img <- pdf_image_new(w, heights)
pdf_draw_image(page, img, 72, 380, 451, 316)

# Paths: a frame round the image and a triangle.
pdf_draw(page, data.frame(op = "rect", x = 72, y = 380, w = 451, h = 316),
  stroke = "grey30")
pdf_draw(page, list(x = c(260, 335, 297.6), y = c(250, 250, 330)),
  fill = "firebrick", close = TRUE)
pdf_page_end(page)

page <- pdf_page_new(w)
pdf_draw_text(page, 72, 760, "Page two", size = 18)
pdf_draw_text(page, 72, 730, "A second page, to have something to rearrange.")
pdf_page_end(page)

report <- file.path(dir, "report.pdf")
pdf_save(w, report)
```

Fonts are the 14 standard PDF fonts by name, or a TrueType or OpenType file
given to `pdf_font()`, which embeds it. `pdf_text_width()` measures text, for
laying out lines: `vignette("writing")` builds a paged report from it.

## Read it back

`pdf_open()` takes a path, a raw vector or a connection. It reads only the
cross-reference table; each function then reads what it needs.

```{r}
pdf <- pdf_open(report)
pdf_meta(pdf)[c("version", "pages", "title", "author")]
pdf_pages(pdf)[, c("page", "width", "height", "rotate")]
```

Text comes back per page, in reading order:

```{r}
cat(pdf_page_text(pdf, pages = 1))
```

Fonts and images. `pdf_image()` returns an image's samples, or, for 8-bit grey
and RGB, a `nativeRaster` R can draw:

```{r, fig.width = 5, fig.height = 3.5, fig.alt = "The volcano image, read back from the PDF"}
pdf_font_table(pdf)
imgs <- pdf_page_images(pdf, 1)
imgs
grey <- pdf_image(pdf, imgs$object[1], as = "native")
dim(grey)
grid::grid.raster(grey, interpolate = FALSE)
```

Underneath, every object is reachable as R values: `pdf_objects()`,
`pdf_object()` and `pdf_stream()`, covered in `vignette("zupdf")`.

## Words and where they are

`pdf_data()` gives each word with its box, in points from the top left, as
pdftools does:

```{r}
head(pdf_data(report)[[1]])
```

## Take it apart, put it together

Splitting, merging, reordering and rotating are all one operation: copying
pages from open files into a new writer. Each copied page brings the fonts and
images it uses.

```{r}
w <- pdf_new()
pdf_copy_pages(w, pdf, pages = 2)               # page two first
pdf_copy_pages(w, pdf, pages = 1, rotate = 90)  # then page one, turned
pdf_set_meta(w, title = "Rearranged")
rearranged <- pdf_open(pdf_save(w))
pdf_pages(rearranged)[, c("page", "rotate")]
pdf_close(rearranged)
```

`vignette("assembling")` has more, including pages from several files.

## Lock it

Encryption is set on the writer before any page is added. Permissions say what
a reader that honours them may do.

```{r}
w <- pdf_new()
pdf_set_encryption(w, user = "open sesame", owner = "admin",
  permissions = c("print", "copy"))
pdf_copy_pages(w, pdf)
locked <- file.path(dir, "locked.pdf")
pdf_save(w, locked)

tryCatch(pdf_open(locked), zupdf_password_error = function(e) "a password is needed")
secret <- pdf_open(locked, password = "open sesame")
pdf_meta(secret)[c("encryption", "permissions")]
pdf_close(secret)
pdf_close(pdf)
```

## The pdftools and qpdf names

Code written for pdftools or qpdf runs with `library(zupdf)` in their place,
for everything but rendering: the same function names, arguments and results.

```{r}
pdf_info(report)$pages
substr(pdf_text(report)[1], 1, 40)
pdf_length(report)

both <- pdf_combine(c(report, report), output = file.path(dir, "both.pdf"))
pdf_length(both)
first <- pdf_subset(both, pages = 1, output = file.path(dir, "first.pdf"))
pdf_length(first)
```

`vignette("switching")` lists what is covered, where results differ, and what
happens when both packages are attached.

## Untrusted files

A PDF from elsewhere may be built to cause harm. `pdf_open()` bounds the input
size, the numbers of objects and pages, the size of any decoded stream and the
depth of nesting. Going past a limit, a damaged file and a wrong password are
each their own condition class, under `zupdf_error`:

```{r}
tryCatch(
  pdf_open(report, max_pages = 1),
  zupdf_limit_error = function(e) class(e)
)
tryCatch(
  pdf_open(charToRaw("not a PDF")),
  zupdf_error = function(e) class(e)
)
```

`?zupdf-conditions` lists them all.
